---
title: "RRactivity"
author: "Noman Khalid"
date: "Sunday, February 15, 2015"
output: html_document
---

##Loading and Processing the Data

```{r, echo=TRUE}
library(lattice)

rractivity <- read.csv("RRactivity.csv", header = TRUE, 
                       colClasses = c("numeric", "character", "numeric"),
                       sep = ",")

head(rractivity)

rractivity$date <- as.Date(rractivity$date, "%Y-%m-%d")

```

##What is mean total number of steps taken per day?

###total steps

```{r, echo=TRUE}
totalsteps <- aggregate(steps ~ date, data=rractivity, sum, na.rm=TRUE)

```

###Histogram

```{r, echo=TRUE}
hist (totalsteps$steps, main = "Daily Total Steps", xlab = "no. of days", col = "blue")

```

###Mean
```{r, echo=TRUE}
mean(totalsteps$steps)
```

###Median
```{r, echo=TRUE}
median(totalsteps$steps)
```

##What is the average daily activity pattern?

###Time Series Plot

```{r, echo=TRUE}
time_series_plot <- tapply (rractivity$steps, rractivity$interval, mean, na.rm = TRUE)

plot (row.names(time_series_plot), time_series_plot, type = "l", 
      xlab = "Every 5 Minute Interval", ylab = "Daily Average",
      main = "Average steps per interval", col = "green")
```

```{r, echo=TRUE}
max_5min_interval <- which.max(time_series_plot)
names(max_5min_interval)
```

##Imputing missing values

```{r, echo=TRUE}
totalNA_rractivity <- sum(is.na(rractivity))
totalNA_rractivity
```

###Filling NAs

```{r, echo=TRUE}
avgsteps <- aggregate(steps ~ interval, data = rractivity, FUN = mean)
fillingNA <- numeric ()
for (i in 1:nrow(rractivity)) {
        obs <- rractivity [i, ]
        if (is.na(obs$steps)) {
                steps <- subset(avgsteps, interval == obs$interval)$steps
        } else {
                steps <- obs$steps
        }
        fillingNA <- c(fillingNA, steps)
}
```

###Creating new dataset with no NAs

```{r, echo=TRUE}
new_rractivity <- rractivity
new_rractivity$steps <- fillingNA 
```

###Histogram of total steps by day

```{r, echo=TRUE}
stepstotalbyday <- aggregate(steps ~ date, data = new_rractivity, sum, na.rm = TRUE)

hist(stepstotalbyday$steps, main = "Total Steps by Day", xlab = "day", col = "blue")
```

###Mean

```{r, echo=TRUE}
mean(stepstotalbyday$steps)
```

###Median
```{r, echo=TRUE}
median(stepstotalbyday$steps)
```

##Differences in activity between weekdays and weekends

I am using the filled dataset for this part (new_rractivity)

```{r, echo=TRUE}
day <- weekdays(new_rractivity$date)
daylevel <- vector()
for (i in 1:nrow(new_rractivity)) {
        if (day[i] == "Saturday") {
                daylevel[i] <- "Weekend"
        } else if (day[i] == "Sunday") {
                daylevel[i] <- "Weekend"
        } else {
                daylevel[i] <- "Weekday"
        }
}
new_rractivity$daylevel <- daylevel
new_rractivity$daylevel <- factor(new_rractivity$daylevel)

stepsByDay <- aggregate(steps ~ interval + daylevel, data = new_rractivity, mean)
names(stepsByDay) <- c("interval", "daylevel", "steps")
```

###Weekend vs Weekday Plot

```{r, echo=TRUE}
xyplot(steps ~ interval | daylevel, stepsByDay, type = "l", layout = c(1, 2), xlab = "Interval", ylab = "Number of steps")
```

The graph shows that during weekdays the number of steps taken are high only at the start of day. While during weekends, there is significant activity across the day
